Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-22 doc/naaclhlt2018-latex/paper.tex
some words…

\title{Title: Something snappy about the SLAM competition}

@@

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

@@

Blah Blah \cite{slam18}

\section{Task Approach}

@@

A user's probability of succeeding on an instance is likely related to their
prior experience with that instance. To capture this, we calculated several
features related to past experience. We encoded the number of times the
current exercise's exact sentence had been seen before by the user. \hl{is this
  right Todd?} We also encoded a set of features recording past experience with
the particular instance word. These features were encoded separately for the
instance token and for the instance root word created by lemmatization.

@@

>>>>

\title{Modeling Second-Language Learning from a Psychological Perspective}

@@

\begin{document}
\maketitle
\begin{abstract}
This is the abstract
\end{abstract}

@@

Educational software that aims to teach people new skills, languages, and
even academic subjects have become increasingly popular.  The wide-spread
deployment of these tools has created interesting opportunities to study
the process of learning in extremely large samples of learners in naturalistic
situations. The Duolingo shared task on Second Lanugage Acquisition Modeling (SLAM)
was a competitive modeling challenge run in early 2018~\cite{slam18}.
The challenge, organized by Duolingo\footnote{http://duolingo.com}, a popular second
language learning app, was
to use log data from thousand of users completing millions of exercises 
predict patterns of future translation mistakes in held-out data.  The data set was
divided into three sets covering students learning to translate English to Spanish (ES$\to$EN),
Spanish to English (ES$\to$EN), and French to English (FR$\to$EN).
This paper reports the approach used by our team which ended in third place 
for the ES$\to$EN data set, second place for the ES$\to$EN data set, and third place
for FR$\to$EN.

Our lab's interest in this competition comes from our emphasis on studying human learning
and memory from a psychological perspective.  While the Duolingo SLAM dataset 
heavily emphasizes language learning (e.g., syntax, grammar, semantics), it also touches on
basic learning and memory phenomena (e.g., retrieval of the translated
word from memory).  The study of the psychological processes supporting second-language 
acquisition has long been a topic of interest in the learning and memory literature~\cite[e.g.,]{Atkinson:1972rm,Atkinson:1972rz,Pavlik:2008rm}.
In fact, much of the classic work on psychology of memory involved aspects of verbal
language learning, which helped to identify a number of important and robust phenomena
in memory including the retention curve and the advantage for spaced
over massed practice~\cite{Ruth:1928aa,Rubin:1996aa,Cepeda:2006aa}.

\hl{ Todd: I want to transition to something about how our approach tried to stick to some
psychological principles but i'm not sure... maybe need to wait until al the later sections
are fleshed out to lay out what the "concept" is behind our approach}.

\hl{One idea is to talk about while the research on learnign and memory is important
it might be hard to graft those principles into datasets like these because of 
subject variability, etc... As a result it is unlikely that straightfoward
cognitive models will work but instead need to combine maching learning methods
that can create much more complex rules and decisions with features that reflect
aspects of what makes something easy or hard to learn.}

\cite{Tubridy:2018aa}
\cite{Tubridy:2018ab}
\cite{Corbett1995bkt}
\cite{Roediger2006te,Anderson1994rif,Mozer:2009cs}
\section{Task Approach}

@@

A user's probability of succeeding on an instance is likely related to their
prior experience with that instance. To capture this, we calculated several
features related to past experience. We encoded the number of times the
current exercise's exact sentence had been seen before by the user. 
We also encoded a set of features recording past experience with
the particular instance word. These features were encoded separately for the
instance token and for the instance root word created by lemmatization.
\cite{Perruchet:1998nw,Chun:1999gt} \hl{chun is good, maybe something other
that PARSER as it is more a statistcal learning model}

@@

\section{Acknowledgments}
This research was supported by NSF grant DRL-1631436 and
seed funds from the NYU Dean for Science.
